The perceptual dimensions of \\ sonority-driven epenthesis

### 2013-09-11 paper.tex
Intro to 4C

As with {\sc Sonority Angle}, this gives rise to a family of constraints {\sc Ident(SonRise)}$<n$, defined similarly to {\sc Ident(SonAngle)}$<n$.

@@

\vspace{-1cm}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-kft-o}/} & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & & \textipa{j@kifto} & 2,4!,5 \\
\hline b. & \ding{43} & \textipa{j@kfito} & 2,3,5 \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\noindent {\bf Remark on conflation.} Notice that all distinctions among the set of rising and level sonority clusters on the one hand, and the distinctions among the set of falling sonority clusters, are conflated. For instance, consider the cluster /mst/. {\sc SonAngle}(m,s), at 2.03, is smaller than {\sc SonAngle}(s,t), at 2.11, yet epenthesis occurs on the right, creating the larger {\sc Sonority Angle}. This is because the distinction between {\sc Sonority Angles} of 2.03 and 2.11 is irrelevant under the given  constraint ranking, since both 2.03 and 2.11 violate the relevant faithfulness constraint {\sc Ident(SonAngle)}$<$1.5. This is possible because I have adopted the position that the {\sc Ident(SonAngle)}$<n$ constraints can be freely ranked.

Adopting the converse strategy --- that {\sc Ident(SonAngle)}$<n$ constraints have a universal ranking --- would not predict [\textipa{ms1t}]. Instead we would have the following situation:

@@

@@

% TODO: Tableau

@@

% TODO: tableau

% TODO: summary

Complete tables illustrating the behaviour of the four idiolects are given in Appendix A. % TODO

@@

In this last part of our case study of Chaha, we turn to the behaviour of quadriconsonantal clusters. These epenthesise as [CC\textipa{1}CC] or [C\textipa{1}CC\textipa{1}C]. 
I will show that the rankings already obtained for CCC clusters predict almost exactly which of the CCCC clusters should display variation and in which direction.

@@

The undominated nature of the constraint *CCC, which we used to trigger epenthesis among the triconsonantal clusters, also triggers epenthesis in quadriconsonantal clusters. Notice that using this as the markedness constraint predicts exactly the two outcomes that are directly observed.

Given the general form of the ranking {\sc Ident(SonAngle)}$<n$ $\gg$ {\sc Align-Left},
the expected behaviour of four-consonant clusters is as follows.

@@

>>>>

As with {\sc Sonority Angle}, this gives rise to a family of constraints {\sc Ident(SonRise)}$<n$, defined similarly to {\sc Ident(SonAngle)}$<$n.

@@

\vspace{-1cm}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-\textnormal{CCC}-o}/} & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & & \textipa{j@\textnormal{C}i\textnormal{CC}o} & 2,4!,5 \\
\hline b. & \ding{43} & \textipa{j@\textnormal{CC}i\textnormal{C}o} & 2,3,5 \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\noindent {\bf Remark on conflation.} Notice that all distinctions among the set of rising and level sonority clusters on the one hand, and the distinctions among the set of falling sonority clusters, are conflated. For instance, consider the cluster /mst/. {\sc SonAngle}(m,s), at 2.03, is smaller than {\sc SonAngle}(s,t), at 2.11, yet epenthesis occurs on the right, creating the larger {\sc Sonority Angle}. This is because the distinction between {\sc Sonority Angles} of 2.03 and 2.11 is irrelevant under the given  constraint ranking, since both 2.03 and 2.11 violate the relevant faithfulness constraint {\sc Ident(SonAngle)}$<$1.5. This is possible because I have adopted the position that the {\sc Ident(SonAngle)}$<$n constraints can be freely ranked.

Adopting the converse strategy --- that {\sc Ident(SonAngle)}$<$n constraints have a universal ranking --- would not predict [\textipa{ms1t}]. Instead we would have the following situation:

@@

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@}-CCC-\textipa{o}/} & {\sc Anchor-Left Rt} \\[0.5ex]
\hline \hline a. & & \textipa{j@.\textnormal{C}1\textnormal{C}.\textnormal{C}o} &  \\
\hline b. & \ding{43} & \textipa{j@\textnormal{C}.\textnormal{C}1.\textnormal{C}o} & $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

@@

In summary, we were able to explain all four CCC idiolects of Chaha by ranking different {\sc Ident(SonAngle)}$<$n constraints higher than either {\sc Align-Left} or {\sc Anchor-Left Root}.

@@

In this last part of our case study of Chaha, we turn to the behaviour of quadriconsonantal clusters. The only such clusters come from the imperative/jussive CCC stem followed by the consonant-initial suffix -\textipa{n@}. The underlying form is /\textipa{n-CCC-n@}/. They surface either as [\textipa{n1\underline{\textnormal{CC}1\textnormal{C}n}@}] or [\textipa{n1\underline{\textnormal{C}1\textnormal{CC}1n}@}].

In this section, I will demonstrate that the rankings already obtained for CCC clusters predict almost exactly which of the CCCC clusters should display variation and in which direction.

@@

The undominated nature of the constraint *CCC, which we used to trigger epenthesis among the triconsonantal clusters, also triggers epenthesis in quadriconsonantal clusters.
This excludes possible solutions such as *[C\textipa{1}CCC] and *[CCC\textipa{1}C], as these still incur a violation of *CCC.

Solutions such as [C\textipa{1}C\textipa{1}CC] and [CC\textipa{1}C\textipa{1}C] are also excluded, as these incur unnecessary violations of {\sc Ident(SonAngle)} and {\sc Dep}. If the middle two consonants C2 and C3 are split up by an epenthetic vowel, there is no further need for epenthesis as *CCC is already satisfied by [CC\textipa{1}CC].

Therefore, the only two possible solutions under my approach are [CC\textipa{1}CC] and [C\textipa{1}CC\textipa{1}C]. Which one is chosen depends on the sonority profile of the clusters, and which of {\sc Align-Left} and {\sc Anchor-Left Root} is dominant.

\ex. {\sc Align-Left} favours [CC\textipa{1}CC], but {\sc Anchor-Left Rt} favours [C\textipa{1}CC\textipa{1}C].

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c:c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{n-\textnormal{CCC}-n@}/} & Align-Left & Anchor-Left Root \\[0.5ex]
\hline \hline a. & & \textipa{n1\textnormal{C.C}1\textnormal{C}.n@} & 2,3,5,6 & $\ast$ \\
\hline b. & \ding{43} & \textipa{n1.\textnormal{C}1\textnormal{C.C}1.n@} & 2,4!,5,7 & \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

There are some clusters whose epenthetic behaviour is subject to other considerations. The constraint {\sc Jussive-R} still applies to these forms, for example, which ensures that the following forms epenthesise as [CC\textipa{1}CC]:

\ex. \a. \textipa{n1-\underline{nk'1m-n}@}
     \b. \textipa{n1-\underline{nf1k-n}@}
     \c. \textipa{n1-\underline{rm1d-n}@}
     \z. \citep[(33a,34a,34b)]{rose.2000}

In addition, Chaha prohibits intramorphemic geminates, which enforces the following forms:

\ex. \a. \textipa{n1-\underline{k'f1fn}@}
     \b. \textipa{n1-\underline{sd1dn}@}
     \z. \citep[(37a,b)]{rose.2000}

I will not consider such clusters further.

Given the general form of the ranking {\sc Ident(SonAngle)}$<$n $\gg$ {\sc Align-Left} $\gg$ {\sc Anchor-Left Root}, the expected behaviour of four-consonant clusters is as follows: all /CCCC/ clusters surface as [CC\textipa{1}CC], except when {\sc SonAngle}(C2,C3) exceeds n and {\sc SonAngle}(C1,C2) and {\sc SonAngle}(C3,C4) are both smaller than n. Illustrative tableaux are shown below for n=1.5:

\ex. TODO tableaux

Given the general form of the ranking {\sc Ident(SonAngle)}$<$n $\gg$ {\sc Anchor-Left Root} $\gg$ {\sc Align-Left}, the default behaviour is to epenthesise as [C\textipa{1}CC\textipa{1}C]. This preference is overridden when at least one of {\sc SonAngle}(C1,C2) and {\sc SonAngle}(C3,C4) are greater than n and {\sc SonAngle}(C2,C3) is less than n.

\ex. TODO tableaux

@@

The following table shows the actual surface forms observed for each of the underlying clusters, as well as the predictions of the four idiolectal rankings.
